---
title: "BN network"
author: "ZHOU WEI"
date: "7/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(visNetwork)
library(bnlearn)
library(rbmn)
library(Rgraphviz)
library(gRain)
```

# Building the model
## Structure learning
```{r}

mishap.dag = empty.graph(nodes = c("QL","WL", "RL", "SA","TL","OS","GN","PL","EE","HF","US","MS"))
# RL: operation rule
# WL: workload
# QL: qualification
# TL: telemetry
# NG: navigation
# PL: power plant
# OS: obstacle sensing
# US: uav
# WT: weather
# HF: human factor
# MS: mishap

mishap.dag = set.arc(mishap.dag, from = "TL", to = "US")
mishap.dag = set.arc(mishap.dag, from = "OS", to = "US")
mishap.dag = set.arc(mishap.dag, from = "PL", to = "US")
mishap.dag = set.arc(mishap.dag, from = "GN", to = "US")
mishap.dag = set.arc(mishap.dag, from = "WL", to = "SA")
mishap.dag = set.arc(mishap.dag, from = "QL", to = "SA")
mishap.dag = set.arc(mishap.dag, from = "RL", to = "HF")
mishap.dag = set.arc(mishap.dag, from = "TL", to = "HF")
mishap.dag = set.arc(mishap.dag, from = "SA", to = "EE")
mishap.dag = set.arc(mishap.dag, from = "SA", to = "HF")
mishap.dag = set.arc(mishap.dag, from = "EE", to = "MS")
mishap.dag = set.arc(mishap.dag, from = "HF", to = "MS")
mishap.dag = set.arc(mishap.dag, from = "US", to = "MS")

arc.set = matrix(c(
  "TL", "US",
  "OS", "US",
  "PL", "US",
  "GN", "US",
  "SA", "EE",
  "WL", "SA",
  "QL", "SA",
  "RL", "HF",
  "TL", "HF",
  "SA", "HF",
  "EE", "MS",
  "HF", "MS",
  "US", "MS"),
  byrow = TRUE, ncol = 2,
  dimnames = list(NULL, c("from", "to")))

arcs(mishap.dag) = arc.set

```


```{r}
mishap.dag
```



## Parameters learning
```{r}
RL.lv = c("BLOVS", "VLOS")
WL.lv = c("danger","warning","suggestion","safe")
QL.lv = c("basic","adequate","skilled")
TL.lv = c("fail","pass")
GN.lv = c("fail","pass")
PL.lv = c("fail","pass")
OS.lv = c("fail","pass")
EE.lv = c("bad","good")
SA.lv = c("lack","pass")
US.lv = c("fail","pass")
HF.lv = c("fail","pass")
MS.lv = c("fail","pass")


RL.prob = array(c(0.5, 0.5), dim = 2, dimnames = list(RL = RL.lv))
WL.prob = array(c(0.4, 0.3, 0.2, 0.1), dim = 4, dimnames = list(WL = WL.lv))
QL.prob = array(c(0.6, 0.3, 0.1), dim = 3, dimnames = list(QL = QL.lv))
TL.prob = array(c(0.004, 0.996), dim = 2, dimnames = list(TL = TL.lv))
GN.prob = array(c(0.001, 0.999), dim = 2, dimnames = list(GN = GN.lv))
PL.prob = array(c(0.006, 0.994), dim = 2, dimnames = list(PL = PL.lv))
OS.prob = array(c(0.001, 0.999), dim = 2, dimnames = list(OS = OS.lv))
EE.prob = array(c(0.964, 0.036, 0.01, 0.99), dim = c(2,2), dimnames = list(EE = EE.lv, SA = SA.lv))

SA.prob = array(c(0.9999, 0.0001,
                  0.999, 0.001,
                  0.75, 0.25,
                  0.4, 0.6,
                  0.999, 0.001,
                  0.99, 0.01,
                  0.7, 0.3,
                  0.35, 0.65,
                  0.99, 0.01,
                  0.9, 0.1,
                  0.65, 0.35, 
                  0.2, 0.8), dim = c(2, 4, 3),
                dimnames = list(SA = SA.lv, WL = WL.lv, QL = QL.lv))

HF.prob = array(c(0.999, 0.001,
                  0.55, 0.45,
                  0.75, 0.25,
                  0.35, 0.65,
                  0.9999, 0.0001,
                  0.85, 0.15,
                  0.95, 0.05,
                  0.38, 0.62),
                dim = c(2, 2, 2, 2),
                dimnames = list(HF = HF.lv, SA = SA.lv, 
                                TL = TL.lv,RL = RL.lv))

US.prob = array(c(1, 0,
                  1, 0,
                  1, 0,
                  1, 0,
                  0.99, 0.01,
                  0.95, 0.05,
                  0.93, 0.07,
                  0.87, 0.13,
                  1, 0,
                  1, 0,
                  1, 0,
                  1, 0,
                  0.98, 0.02,
                  0.72, 0.28,
                  0.65, 0.35,
                  0, 1), dim = c(2, 2, 2, 2, 2),
                dimnames = list(US = US.lv, TL = TL.lv,
                                OS = OS.lv, PL = PL.lv,
                                GN = GN.lv))


MS.prob = array(c(1, 0,
                  0.9999, 0.0001,
                  1, 0,
                  0.999, 0.001,
                  0.9998, 0.0002,
                  0.02, 0.98,
                  0.35, 0.65,
                  0, 1),
                dim = c(2,2,2,2),
                dimnames = list(MS = MS.lv, HF = HF.lv,
                                EE = EE.lv, US = US.lv))


CPT = list(WL = WL.prob, QL = QL.prob,
           RL = RL.prob, SA = SA.prob,
           EE = EE.prob, HF = HF.prob,
           TL = TL.prob, GN = GN.prob,
           PL = PL.prob, OS = OS.prob,
           US = US.prob, MS = MS.prob)

```

```{r}
CPT
```

# Inference 

```{r}
mishap = custom.fit(mishap.dag,CPT)
```


## model visualization

```{r,fig.height=10, fig.width=12}

graphviz.chart(mishap, type = "barprob", 
               grid = TRUE, 
               bar.col = "darkgreen",
               strip.bg = "lightskyblue",
               scale = c(3,3),
               draw.levels = TRUE)

```

## Hightlight the Human Factor Module
```{r}

hlight <- list(nodes = nodes(mishap.dag), arcs = arcs(mishap.dag),
               col = "black", textCol = "black")
pp <- graphviz.plot(mishap.dag, 
                    highlight = hlight, shape = "ellipse",
                    layout = "dot",
                    groups = list(c("WL","QL","RL","TL","OS","PL","GN"),
                                  c("SA"),
                                  c("EE","HF","US"),
                                  c("MS")))

edgeRenderInfo(pp) <- list(
  col = c("WL~SA" = "darkblue", "QL~SA" = "darkblue",
  "SA~HF" = "darkblue","RL~HF" = "darkblue", 
  "TL~HF" = "darkblue","EE~HF" = "darkblue"),
  lwd = c("WL~SA" = 3, "QL~SA" = 3,
  "SA~HF" = 3,
  "RL~HF" = 3, "TL~HF" = 3),
  lty = c("WL~SA" = "dashed", "QL~SA" = "dashed",
  "SA~HF" = "dashed", 
  "RL~HF" ="dashed", "TL~HF" = "dashed"))

nodeRenderInfo(pp) <- list(fill = c("HF" = "#BEBADA",
                                   "RL" = "#BEBADA",
                                   "SA" = "#BEBADA",
                                   "WL" = "#BEBADA",
                                   "QL" = "#BEBADA"))

renderGraph(pp)
```


## Hightlight the UAV Module
```{r}
hlight <- list(nodes = nodes(mishap.dag), arcs = arcs(mishap.dag),
               col = "black", textCol = "black")
pp <- graphviz.plot(mishap.dag, 
                    highlight = hlight, shape = "ellipse",
                    layout = "dot",
                    groups = list(c("WL","QL","RL","TL","OS","PL","GN"),
                                  c("SA"),
                                  c("EE","HF","US"),
                                  c("MS")))

edgeRenderInfo(pp) <- list(
  col = c("TL~US" = "darkblue", "GN~US" = "darkblue",
  "OS~US" = "darkblue","PL~US" = "darkblue"),
  lwd = c("TL~US" = 3, "GN~US" = 3,
  "OS~US" = 3,"PL~US" = 3),
  lty = c("TL~US" = "dashed", "GN~US" = "dashed",
  "OS~US" = "dashed","PL~US" = "dashed"))

nodeRenderInfo(pp) <- list(fill = c("PL" = "#8DD3C7",
                                   "OS" = "#8DD3C7",
                                   "GN" = "#8DD3C7",
                                   "TL" = "#8DD3C7",
                                   "US" = "#8DD3C7"))

renderGraph(pp)
```

## Hightlight the External Environment Module
```{r}
hlight <- list(nodes = nodes(mishap.dag), arcs = arcs(mishap.dag),
               col = "black", textCol = "black")
pp <- graphviz.plot(mishap.dag, 
                    highlight = hlight, shape = "ellipse",
                    layout = "dot",
                    groups = list(c("WL","QL","RL","TL","OS","PL","GN"),
                                  c("SA"),
                                  c("EE","HF","US"),
                                  c("MS")))

edgeRenderInfo(pp) <- list(
  col = c("SA~EE" = "darkblue"),
  lwd = c("SA~EE" = 3),
  lty = c("SA~EE" = "dashed"))

nodeRenderInfo(pp) <- list(fill = c("EE" = "#FDB462"))

renderGraph(pp)
```


## Posterior for each single input variable

```{r}
bn.fit.barchart(mishap$WL, xlab = "Probabilities of Fatigue", 
                ylab = "Levels", main = "Workload")
bn.fit.barchart(mishap$QL, xlab = "Probabilities of Inexperience", 
                ylab = "Levels",  main = "Qualification")
bn.fit.barchart(mishap$SA, 
                xlab = "P(SA|WL,QL)", 
                ylab = " ",
                main = "Situational Awareness")
bn.fit.barchart(mishap$EE, xlab = "P(EE|SA)", 
                ylab = " ",
                main = "External Environment")

```

## Generate random samples

```{r}

rbn(mishap,n=10)
write.csv(rbn(mishap,n=20),"random.csv")

```


## Compute probabilities with Maximum a Posteriori queries

- Under BVLOS, basic qualification, danger workload  
good weather

```{r,fig.height=10, fig.width=12}

EE.1 = array(c(0, 1, 0, 1), dim = c(2,2), dimnames = list(EE = EE.lv, SA = SA.lv))
QL.1 = array(c(1,0,0), dim = 3, dimnames = list(QL = QL.lv))
WL.1 = array(c(1, 0, 0, 0), dim = 4, dimnames = list(WL = WL.lv))
RL.1 = array(c(0, 1), dim = 2, dimnames = list(RL = RL.lv))


CPT.1 = list( TL = TL.prob, GN = GN.prob,PL = PL.prob, 
              OS = OS.prob, RL = RL.1, WL = WL.1, 
              QL = QL.1,SA = SA.prob, US = US.prob,
              HF = HF.prob, EE = EE.1, MS = MS.prob)

mishap.1 = custom.fit(mishap.dag,CPT.1)

graphviz.chart(mishap.1, type = "barprob", 
               grid = TRUE, 
               bar.col = "darkorange1",
               strip.bg = "lightpink",
               scale = c(3,3),
               layout = "dot")
```



